---
title: "Euro 2021 Predictor"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Gillard BT"
date: "Updated: `r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    theme: cerulean
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)
```

```{r FormatData, message=FALSE, warning=FALSE}
library(tidyverse)
#read in real life results
TrueResults<- readxl::read_excel("EuroPredictorTrueResults.xlsm")
TrueResults<- TrueResults[7:42, 3:6]
colnames(TrueResults)<- c("Home", "HomeScore", "AwayScore", "Away")
TrueResults<- cbind(TrueResults, Result = ifelse(TrueResults$HomeScore == TrueResults$AwayScore, "Draw", ifelse(TrueResults$HomeScore > TrueResults$AwayScore, TrueResults$Home, TrueResults$Away))) #result information
AllPreds<- list(CorrectResults = TrueResults) #seed list so player predictions can be added

#get predictions from every player (use loop calling all files in Entries directory)
entries<- list.files("Entries/", full.names = TRUE)
for(file in entries){
predictionsRoot<- readxl::read_excel(file) #change to file
player<- gsub(" ", "", as.character(predictionsRoot[47,2]),  fixed = TRUE)
predictions<- predictionsRoot[7:42, 3:6]
colnames(predictions)<- c("Home", "HomeScore", "AwayScore", "Away")
predictions<- cbind(predictions, Result = ifelse(predictions$HomeScore == predictions$AwayScore, "Draw", ifelse(predictions$HomeScore > predictions$AwayScore, predictions$Home, predictions$Away))) #result information
AllPreds<- append(AllPreds, list(predictions)) #append player's predictions to list seeded by correct results
names(AllPreds)[length(AllPreds)]<- player #rename list with player name
}

Results<- AllPreds #Renaming because I'd originally saved and reimported results so all code below is written for the results object

#date information
ResultsSoFar<- readxl::read_excel("EuroPredictorTrueResults.xlsm")[7:42, c(1,3:6)]
ResultsSoFar$`EURO 2020 PREDICTOR`<- gsub(x = ResultsSoFar$`EURO 2020 PREDICTOR`, pattern = "..., ", replacement = "") %>% gsub(pattern = "June ", replacement = "06-") %>% as.Date(format = "%m-%d") #get rid of days and reformat
ResultsSoFar<- ResultsSoFar[order(ResultsSoFar$`EURO 2020 PREDICTOR`),]
library(kableExtra)
ResultsSoFar[!is.na(ResultsSoFar[,3]),] %>% kbl(caption = "Euro 2020 Results so far", align = "c", row.names = FALSE, col.names = NULL, linesep = "") %>% kable_classic(html_font = "Cambria") %>% kable_styling("striped", font_size = 10)
```


```{r ImportResults}
#loop over each row?
  #cbind points to each player? with lapply?
invisible(unlist(lapply(names(Results[2:length(Results)]), function(x) identical(Results[[1]][[1]], Results[[x]][[1]])))) #check format is identical by checking home team names

#successful points vector for one player (first player in the Results list object)
#unlist(lapply(1:nrow(Results$CorrectResults[!is.na(Results$CorrectResults$HomeScore),]), function(row) ifelse(identical(Results[[1]][[row,2]], Results[[2]][[row,2]]) & identical(Results[[1]][[row,3]], Results[[2]][[row,3]]), 3, ifelse(identical(Results[[1]][[row,5]], Results[[2]][[row,5]]), 1, 0))))

Scores<- lapply(names(Results[2:length(Results)]), function(x) unlist(lapply(rownames(Results$CorrectResults[!is.na(Results$CorrectResults$HomeScore),]), function(row) ifelse(identical(Results[[1]][[row,2]], Results[[x]][[row,2]]) & identical(Results[[1]][[row,3]], Results[[x]][[row,3]]), 3, ifelse(identical(Results[[1]][[row,5]], Results[[x]][[row,5]]), 1, 0)))))

names(Scores)<- names(Results[2:length(Results)])
```

**Welcome to the EURO 2020 Predictor Championship!** The aim of the game is simple, predict the scores for all of the group stage games at this year's tournament. Each correct *score* earns 3 points. Each correct *result* earns 1 point.

We have **`r length(Results)-2` players** involved so we have a healthy kitty to pay out for the **top 5 players. £`r round(x = (length(Results)-2)*5*0.45, digits = 2)` goes to the top dog. £`r round(x = (length(Results)-2)*5*0.25, digits = 2)` for the top loser. £`r round(x = (length(Results)-2)*5*0.15, digits = 2)` for the best of the rest. And £`r round(x = (length(Results)-2)*5*0.075, digits = 2)` for 4th and 5th places.**

On the [Leaderboard] you'll see an extra player called *RandomControlEntry* in <span style="color: red;">red</span>. These predictions were generated randomly using the macro included in the spreadsheet. So if you have fewer points than *RandomControlEntry*, any thought you put into your entry was an utter waste of time. 

Go to [Results Tracker] for a more in depth look at which games scored you points.

#Leaderboard

```{r Leaderboard, message=FALSE, warning=FALSE}
LeaderBoard<- data.frame(Player = names(Scores), Points = unlist(lapply(Scores, sum)), CorrectScore = unlist(lapply(Scores, function(x) sum(x == 3))), CorrectResult = unlist(lapply(Scores, function(x) sum(x == 1))), row.names = NULL)

LeaderBoard<- LeaderBoard[order(LeaderBoard$Points, LeaderBoard$CorrectScore, decreasing = TRUE),]
rownames(LeaderBoard)<- 1:nrow(LeaderBoard)

kbl(LeaderBoard, caption = "Table of Champions (Coloured rows indicate money winners)", align = "l", row.names = TRUE) %>% kable_classic(html_font = "Cambria") %>%  row_spec(1, bold = T, background = "#c9b037") %>% row_spec(2, bold = T, background = "#b4b4b4") %>% row_spec(3, bold = T, background = "#ad8a56") %>% row_spec(4:5, bold = T, background = "#3CB371") %>% row_spec(which(LeaderBoard$Player == "RandomControlEntry"), italic = T, color = "red")
```

> When 2 players have the same points tally, correct scores are taken into account. If players are still equal, the tie break question "How many total goals (including knockout games but excluding penalty shoot outs) will England score at Euro 2020?" will come into play. Closest wins.

##England Goals

Tracks the tie break question of how many goals England will score. The TrueGoals are updated after every England game.

```{r englandGoals, message=FALSE, warning=FALSE}
par(mar=c(7,4,4,4))
TrueGoals<- as.numeric(readxl::read_excel("EuroPredictorTrueResults.xlsm")[44,7])
AllGoals<- list(TrueGoals = TrueGoals) #seed list so player predictions can be added
for(file in entries){
predictionsRoot<- readxl::read_excel(file) #change to file
player<- gsub(" ", "", as.character(predictionsRoot[47,2]),  fixed = TRUE)
EngGoals<- as.numeric(predictionsRoot[44,7])
#result information
AllGoals<- append(AllGoals, list(EngGoals)) #append player's predictions to list seeded by correct results
names(AllGoals)[length(AllGoals)]<- player #rename list with player name
}
barplot(unlist(AllGoals), main = "Tie Breaker: England Goals scored", col = c("red", "white"), las = 2, cex.names = 0.7)
abline(h = AllGoals$TrueGoals, lty = 2, col = "red")
```

#Hypothetical Euros

Who do we, as a collective of Mystic Megs, predict will be the winners and losers in the group stage? This plot shows how many times each team has been predicted to win a match in the group stages.

```{r IfwePredicted, fig.show='hold'}
WinnerCount<- sapply(names(Results[2:length(Results)]), function(x) summary(as.factor(Results[[x]][["Result"]])), USE.NAMES = TRUE)

barplot(sort(sapply(unique(Results$CorrectResults$Home), function (x) sum(grepl(x = names(unlist(WinnerCount)), pattern = x)))), horiz = TRUE, las = 1, cex.names = 0.5, col = viridis::rocket(24), main = "Total (all players combined) predicted Wins for each team", xlab = "Number of wins predicted", xlim=range(pretty(c(0, max(sapply(unique(Results$CorrectResults$Home), function (x) sum(grepl(x = names(unlist(WinnerCount)), pattern = x)))+1)))))

#top predicted result
#barplot(table(unlist(lapply(names(WinnerCount), function(x) names(WinnerCount[[x]])[which(WinnerCount[[x]] == max(WinnerCount[[x]]))]))), horiz = TRUE, las = 1, cex.names = 0.5, col = viridis::rocket(10), main = "Overall most predicted results", xlab = "Number of wins predicted", xlim=range(pretty(c(0, 10))))

#bottom predicted result
#barplot(table(unlist(lapply(names(WinnerCount), function(x) names(WinnerCount[[x]])[which(WinnerCount[[x]] == min(WinnerCount[[x]]))]))), horiz = TRUE, las = 1, cex.names = 0.5, col = viridis::magma(10), main = "Overall least predicted results", xlab = "Number of wins predicted", xlim=range(pretty(c(0, 3), n = 3)))
```

More specifically, how many times does each player think the home nations will win in the group stage  (more teams available on request)? And how many boring draws do people think we'll suffer through?

```{r specificTeams, fig.show='hold', warning=FALSE}
par(mar=c(7,4,4,4))
#england wins
barplot(sapply(names(Results[1:length(Results)]), function(x) sum(Results[[x]][["Result"]] == "England"), USE.NAMES = TRUE), main = "Predicted number of England Wins", col = c("red", "white"), las = 2, cex.names = 0.7)
abline(h = sum(Results$CorrectResults[["Result"]] == "England"), lty = 2, col = "black")

#scotland wins
barplot(sapply(names(Results[1:length(Results)]), function(x) sum(Results[[x]][["Result"]] == "Scotland"), USE.NAMES = TRUE), main = "Predicted number of Scotland Wins", col = c("blue", "white"), las = 2, cex.names = 0.7)
abline(h = sum(Results$CorrectResults[["Result"]] == "Scotland"), lty = 2, col = "black")

#Wales wins
barplot(sapply(names(Results[1:length(Results)]), function(x) sum(Results[[x]][["Result"]] == "Wales"), USE.NAMES = TRUE), main = "Predicted number of Wales Wins", col = c("red", "green"), las = 2, cex.names = 0.7)
abline(h = sum(Results$CorrectResults[["Result"]] == "Wales"), lty = 2, col = "black")

#Poland wins
barplot(sapply(names(Results[1:length(Results)]), function(x) sum(Results[[x]][["Result"]] == "Poland"), USE.NAMES = TRUE), main = "Predicted number of Poland Wins", col = c("white", "red"), las = 2, cex.names = 0.7)
abline(h = sum(Results$CorrectResults[["Result"]] == "Poland"), lty = 2, col = "black")

#draws
barplot(sapply(names(Results[1:length(Results)]), function(x) sum(Results[[x]][["Result"]] == "Draw"), USE.NAMES = TRUE), main = "Predicted number of Draws", col = viridis::turbo(length(Results)), las = 2, cex.names = 0.7)
abline(h = sum(Results$CorrectResults[["Result"]] == "Draw"), lty = 2, col = "black")
```


##Results Tracker

This Results Tracker has a bit more depth than the Leaderboard. It shows what points everyone has scored for each game.

Hover over each tile for more information (might not work on phones). Green = Correct Score (3 points), Yellow = Correct Result (1 point), Grey = Wrong! (nil pois).

```{r indepthresults, message=FALSE, warning=FALSE}
#create matrix with points
#ResultsMat<- ResultsSoFar #use date ordered fixtures
ResultsMat<- data.frame(cbind(Game = paste(Results$CorrectResults$Home, "-", Results$CorrectResults$Away))) #create a column combining the match teams into one
#ResultsMat<- ResultsMat[,-c(2:5)] #remove useless info

#points need to be recalculated to include NAs (fixtures yet to be played)
TrackerScores<- lapply(names(Results[2:length(Results)]), function(x) unlist(lapply(1:nrow(Results$CorrectResults[Results$CorrectResults$HomeScore,]), function(row) ifelse(identical(Results[[1]][[row,2]], Results[[x]][[row,2]]) & identical(Results[[1]][[row,3]], Results[[x]][[row,3]]), 3, ifelse(identical(Results[[1]][[row,5]], Results[[x]][[row,5]]), 1, ifelse(is.na(Results[[1]][[row,5]]), NA, 0))))))
names(TrackerScores)<- names(Results[2:length(Results)]) #rename using player names

for(P in names(TrackerScores)){
ResultsMat<- cbind(ResultsMat, TrackerScores[[P]])
} #bind all results (could be adapted to use apply functions)
colnames(ResultsMat)<- c("Game", names(TrackerScores))
rownames(ResultsMat)<- ResultsMat$Game
ResultsMat[1]<- NULL #for ease of plotting, change match descriptor to rownames then remove "Game" column


library(heatmaply) #interactive heatmap with hover information
heatmaply(ResultsMat, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        grid_color = "white",
        grid_size = 0.0001,
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(ResultsMat),
        labRow = ResultsMat$Game,
        colors = c("lightgrey", "yellow", "green"),
        label_names = c("Game", "Player", "Points"),
        )
```

**If I've got your number, you will already have been added to the Euro Predictor Whatsapp group. If you haven't been added, please let me know and I'll send you a joining link. It's the best place to gloat/cry about your guesses.**
